Viewing Computer Science through Citation Analysis

### 2020-03-30 Scientometrics/rebuttal/comp_revised.tex
resubmit

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestorew
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{listings}
%\usepackage{comment}
\usepackage{hyperref}
\usepackage{changes}
\definechangesauthor[name={George Chacko}, color=orange]{gc}
%\setremarkmarkup{(#2)}
%\usepackage[firstpage]{draftwatermark}

@@

\begin{abstract}w
Computer science has experienced dramatic growth and diversification over the last twenty years. Towards a current understanding of the structure of this discipline, we analyze 
a \replaced{large sample}{cohort} of the computer science literature from the DBLP database. For insight on the features of this cohort and the relationship within its components, we have constructed article 
level clusters based on either direct citations or co-citations, and reconciled them \replaced{with}{to} major and minor subject categories in the All Science Journal Classification (ASJC). 
We describe complementary insights from clustering by direct citation and co-citation. \deleted[remark=repositioned]{, and both point to the increase in computer science publications and their scope}.
Our analysis \replaced{reveals}{shows} cross-category clusters, some that interact with external fields, such as the biological sciences, while others remain inward looking.\added{Overall, we document an increase in computer science publications and their scope.}

@@

We XXX XXXX XXXXX

@@

\added{We began by identifying highly cited articles in \emph{comp} and selected those in the 90th percentile (212,311 articles). We then identified ~4.3 million publications in Scopus that cite these 212,311 articles. For each of the 4.3 million citing publications in turn, all possible ${n \choose 2}$ reference pairs were generated from a publication's cited references, where $n$ is the number of references in a publication. The cited reference pairs this generated were then restricted to those where both members of a pair were in the set of 212,311 highly cited papers previously identified. A total of 46,463,117 unique co-cited pairs were thus obtained. The frequency of these co-cited pairs was then computed across the \emph{comp} dataset  and normalized using Salton's cosine formula~\cite{salton_citation_1979} to limit dominance by areas with high citation activity. These data were represented in a graph where each node was a publication and the weighted edge between the pair was the normalized co-citation frequency. \par }

@@

 In our high-level study of the structure of a twenty year sample of the computer science literature, we chose to use both traditional journal-based and article-based approaches. Considering that traditional disciplines `may only partly reflect the actual organization of today's scientific research'~\cite{waltman_new_2012}, we constructed article clusters at high levels of aggregation using citations to examine the  computer science literature and also mapped these clusters to journal-based categories to take advantage of both article level and journal level approaches.

@@

At the threshold of 15\%, only 17 of 20 co-citation clusters mapped to at least one minor subject areas. At the 15\% threshold, no co-citation cluster maps to more than three minor subject areas, in contrast to a maximum of eight minor subject areas for clustering by direct citation (Fig. 3, Cluster 17). The threshold had to be reduced to 10\% for all 20 co-citation clusters to map to at least one minor subject area. We interpret these results as indicative of broader clusters created by weaker linkages that accumulate during the agglomerative clustering phase~\cite{small_clustering_1985}.

Clustering by co-citation begins, for each cluster, with a pair of nodes that nucleates its subsequent formation. Thus, we also designated the pair of nodes in a cluster with the strongest edge as its nucleating pair and labeled the cluster by manually labeling the nucleating pair of documents.  
These labels show a high degree of correspondence  to the ASJC minor subject area that accounts for the largest fraction of the nodes in a co-citation cluster (Table~\ref{tab:centroid_reconcile}). 
%The extent of this correspondence %between the manually applied label and the ASJC minor subject area 
Thus, using the nucleating pair as the basis for labeling co-citation clusters (as we generated them) may be valid, although its usefulness is likely to vary according to the data being examined; we also note that manual annotation (although beneficial) is not scalable. We provide the DOIs of these nucleating pairs matched to manually assigned labels for independent review (Table~\ref{tab:centroid}).

@@

The focus of this article was on high-level features, and a preference for simplicity and intuitiveness in the choice of methods. This preference may address questions  regarding generalizability and the specific choices we made, such as the (i) use of the DBLP dataset matched to Scopus, which may not capture all aspects of computer science and its interactions with other fields, (ii) the basis for clustering, and (iii) mapping the results of this clustering against a classification designed around journals rather than individual articles. We believe that an approach that combines journal-level  with article-level analyses is useful for studies of this kind. \added{However, we acknowledge that our approach would not detect emerging fields with few publications that could be important signs of innovation.}

@@

\begin{figure}[ht]
\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.6]{graclus_cocit_fig.pdf}
% figure caption is below the figure
\caption{Intersections between clusters generated using direct citation and co-citation features. x-axis: Clusters generated by Graclus number 0-19. y-axis: Cluster generated by variable level clustering and agglomerative clustering using a modification of Small and Sweeney (1985)~\cite{small_clustering_1985}. Point size (perc) is the percentage of a co-cited cluster that maps to a corresponding Graclus cluster. A minimum threshold of 15\% was set. Graclus cluster 19, the 20th cluster, did not map to any cluster on the y:axis.}
\label{fig:graclus_cocit_fig}       % Give a unique label
\end{figure}
\newpage
\clearpage
\section*{Tables}

\begin{table}[ht]
\caption{Characterization of the \emph{comp} dataset relative to Scopus ASJC minor subject areas.}
\label{tab:comp}       
\begin{tabular}{lrccc}
\hline\noalign{\smallskip}
minor\_subject\_area & Percent of Publications \\
\noalign{\smallskip}\hline\noalign{\smallskip}
Software & 30.1 \\
Computer Science Applications & 25.0 \\ 
Computer Networks and Communications & 21.5 \\
Theoretical Computer Science & 19.6 \\ 
Computer Science(all) & 19.4 \\ 
Artificial Intelligence & 14.2 \\ 
Information Systems & 11.8 \\ 
Hardware and Architecture & 11.8 \\ 
Signal Processing & 9.8 \\ 
Computer Vision and Pattern Recognition & 9.4 \\ 
Computational Theory and Mathematics & 7.9 \\ 
Human-Computer Interaction & 7.4 \\ 
Computer Graphics and Computer-Aided Design & 5.2 \\ 
Computer Science (miscellaneous) & 0.9 \\ 
\noalign{\smallskip}\hline
\end{tabular}
\end{table}
\newpage

@@

\begin{table}[ht]
\caption{\emph{Nucleating Co-citations.} For each of the 20 co-citation clusters, the pair with the strongest edge (greatest normalized co-citation frequency) is shown below along with a manually assigned label.}
\label{tab:centroid}       
\begin{tabular}{p{8 mm}p{42 mm}ll}
\hline\noalign{\smallskip}
Cluster & Nucleating Pair & NCF & Manual Label \\
\noalign{\smallskip}\hline
18670 & 10.1016/j.cam.2015.03.057 10.1016/j.sigpro.2015.10.009 & 0.92 & Dynamical Systems \\
18892 & 10.1287/msom.1080.0228 10.1287/msom.1060.0190 & 0.72 & Operations Research\\ 
18918 & 10.1007/s11128-010-0177-y 10.1007/s11128-013-0567-z & 0.83 & Image Processing\\ 
18947 & 10.1109/TASE.2011.2160452 10.1109/TASE.2011.2178023 & 0.77 & Robotics\\ 
18956 & 10.1109/ICCV.2017.32 10.1109/ICCV.2017.31 & 0.84 & Computer Vision\\ 
18972 & 10.1109/ASE.2013.6693094 10.1145/2568225.2568254 & 0.69 & Software \\ 
18974 & 10.1016/j.amc.2009.03.023 10.1016/j.amc.2010.07.064 & 0.54 & Applied Mathematics\\ 
18983 & 10.1137/110848864 10.1137/110848876 & 0.66 & Optimization\\ 
18997 & 10.1504/IJMIC.2014.065339 10.1504/IJMIC.2015.068871 & 0.91 & Chaotic Systems \\ 
19008 & 10.1016/j.chb.2008.12.013 10.1016/j.chb.2008.12.012 & 0.63 & Hum Comp Interaction\\ 
19009 & 10.1109/AVSS.2017.8078491 10.1109/ICCV.2017.206 & 0.78 & Artificial Intelligence\\ 
19013 & 10.1145/2508859.2516668 10.1007/978-3-642-42045-0\_15 & 0.88 & Security\\ 
19026 & 10.1145/2541940.2541942 10.1109/HPCA.2014.6835965 & 0.81 & Architecture\\ 
19027 & 10.1016/S0305-0548(03)00250-8 10.1016/j.ejor.2006.03.013 & 0.82 & Graph algorithms\\ 
19030 & 10.1109/ICDE.2008.4497474 10.14778/1687627.1687666 & 0.69 & Databases \\ 
19031 & 10.1109/TMM.2005.843347 10.1109/TCE.2005.1405724 & 0.55 & Networks\\ 
19033 & 10.1109/TIFS.2014.2327757 10.1109/TCYB.2014.2376934 & 0.78 & Facial Recognition\\ 
19034 & 10.1016/j.neuroimage.2013.05.018 10.1016/j.neuroimage.2014.06.016 & 0.63 & Neurology\\ 
19035 & 10.1002/int.21933 10.1002/int.21927 & 1.09 & Intelligent Systems\\ 
19036 & 10.1006/jsco.2000.0402 10.1006/jsco.2000.0403 & 0.73 & Comp Geometry\\ 
\noalign{\smallskip}\hline
\end{tabular}
\end{table}
\newpage

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestorew
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{listings}
%\usepackage{comment}
\usepackage{hyperref}
\usepackage[final]{changes}
%\usepackage{changes}
\definechangesauthor[name={George Chacko}, color=orange]{gc}
%\setremarkmarkup{(#2)}
%\usepackage[firstpage]{draftwatermark}

@@

\begin{abstract}
Computer science has experienced dramatic growth and diversification over the last twenty years. Towards a current understanding of the structure of this discipline, we analyze 
a \replaced{large sample}{cohort} of the computer science literature from the DBLP database. For insight on the features of this cohort and the relationship within its components, we have constructed article 
level clusters based on either direct citations or co-citations, and reconciled them \replaced{with}{to} major and minor subject categories in the All Science Journal Classification (ASJC). 
We describe complementary insights from clustering by direct citation and co-citation. \deleted[remark=repositioned]{, and both point to the increase in computer science publications and their scope}.
Our analysis \replaced{reveals}{shows} cross-category clusters, some that interact with external fields, such as the biological sciences, while others remain inward looking.\added{Overall, we document an increase in computer science publications and their scope.}

@@

\added{We constructed clusters of articles using either direct citation or co-citations and match these clusters to ASJC categories to place them in context. We also align clusters from direct citations to those from co-citations and discuss our findings, Overall, we note cross category clusters with different degrees of interaction with fields outside computer science.}

@@

\added{We began by identifying highly cited articles in \emph{comp} and selected those in the 90th percentile (212,311 articles). We then identified ~4.3 million publications in Scopus that cite these 212,311 articles. For each of the 4.3 million citing publications in turn, all possible ${n \choose 2}$ reference pairs were generated from a publication's cited references, where $n$ is the number of references in a publication. The cited reference pairs generated were then restricted to those where both members of a pair were in the set of 212,311 highly cited papers previously identified. A total of 46,463,117 unique co-cited pairs were thus obtained. The frequency of these co-cited pairs was then computed across the \emph{comp} dataset  and normalized using Salton's cosine formula~\cite{salton_citation_1979} to limit dominance by areas with high citation activity. These data were represented in a graph where each node was a publication and the weighted edge between the pair was the normalized co-citation frequency. \par }

@@

 In our high-level study of the structure of a twenty year sample of the computer science literature, we chose to use both traditional journal-based and article-based approaches. Considering that traditional disciplinary categories `may only partly reflect the actual organization of today's scientific research'~\cite{waltman_new_2012}, we constructed article clusters at high levels of aggregation using citations to examine the  computer science literature and also mapped these clusters to journal-based categories to take advantage of both article level and journal level approaches.

@@

At the threshold of 15\%, only 17 of 20 co-citation clusters mapped to at least one minor subject areas. At the 15\% threshold, no co-citation cluster maps to more than three minor subject areas, in contrast to a maximum of eight minor subject areas for clustering by direct citation \added{(Fig. 3, Cluster 17).} The threshold had to be reduced to 10\% for all 20 co-citation clusters to map to at least one minor subject area. We interpret these results as indicative of broader clusters created by weaker linkages that accumulate during the agglomerative clustering phase~\cite{small_clustering_1985}.

Clustering by co-citation begins, for each cluster, with a pair of nodes that nucleates its subsequent formation. Thus, we also designated the pair of nodes in a cluster with the strongest edge as its nucleating pair and labeled the cluster by manually labeling the nucleating pair of documents.  
These labels show a high degree of correspondence  to the ASJC minor subject area that accounts for the largest fraction of the nodes in a co-citation cluster (Table~\ref{tab:centroid_reconcile}). While using the nucleating pair as the basis for labeling co-citation clusters (as we generated them) may be valid, its usefulness is likely to vary according to the data being examined; we also note that manual annotation (although beneficial) is not scalable. We provide the DOIs of these nucleating pairs matched to manually assigned labels for independent review (Table~\ref{tab:centroid}).

@@

The focus of this article was on high-level features, and a preference for simplicity and intuitiveness in the choice of methods. This preference influences the specific choices we made, such as the (i) use of the DBLP dataset matched to Scopus, which may not capture all aspects of computer science and its interactions with other fields, (ii) the basis for clustering, and (iii) mapping the results of this clustering against a classification designed around journals rather than individual articles. We believe that an approach that combines journal-level  with article-level analyses is useful for studies of this kind. \added{However, we acknowledge that our approach would not detect emerging fields with few publications that could be important signs of innovation. We also feel strongly that contextualized interpretation by domain experts should be performed on results generated by such approaches.}

@@

\begin{figure}[ht]
\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.6]{graclus_cocit_fig.pdf}
% figure caption is below the figure
\caption{Intersections between clusters generated using direct citation and co-citation features. x-axis: Clusters generated by Graclus number 0-19. y-axis: Cluster generated by variable level clustering and agglomerative clustering using a modification of Small and Sweeney (1985)~\cite{small_clustering_1985}. Point size (perc) is the percentage of a co-cited cluster that maps to a corresponding Graclus cluster. A minimum threshold of 15\% was set. Graclus cluster 19, the 20th cluster, did not map to any cluster on the y:axis.}
\label{fig:graclus_cocit_fig}       % Give a unique label
\end{figure}
\newpage
\clearpage
\section{Tables}

\begin{table}[ht]
\caption{Characterization of the \emph{comp} dataset relative to Scopus ASJC minor subject areas under computer science.}
\label{tab:comp}       
\begin{tabular}{lrccc}
\hline\noalign{\smallskip}
minor\_subject\_area & Percent of Publications \\
\noalign{\smallskip}\hline\noalign{\smallskip}
Software & 30.1 \\
Computer Science Applications & 25.0 \\ 
Computer Networks and Communications & 21.5 \\
Theoretical Computer Science & 19.6 \\ 
Computer Science(all) & 19.4 \\ 
Artificial Intelligence & 14.2 \\ 
Information Systems & 11.8 \\ 
Hardware and Architecture & 11.8 \\ 
Signal Processing & 9.8 \\ 
Computer Vision and Pattern Recognition & 9.4 \\ 
Computational Theory and Mathematics & 7.9 \\ 
Human-Computer Interaction & 7.4 \\ 
Computer Graphics and Computer-Aided Design & 5.2 \\ 
Computer Science (miscellaneous) & 0.9 \\ 
\noalign{\smallskip}\hline
\end{tabular}
\end{table}
\newpage

@@

\begin{table}[ht]
\caption{\emph{Nucleating Co-citations.} For each of the 20 co-citation clusters, the pair with the strongest edge (greatest normalized co-citation frequency) is shown below along with a manually assigned label.}
\label{tab:centroid}       
\begin{tabular}{p{8 mm}p{42 mm}ll}
\hline\noalign{\smallskip}
Cluster & Nucleating Pair & NCF & Label \\
\noalign{\smallskip}\hline
18670 & 10.1016/j.cam.2015.03.057 10.1016/j.sigpro.2015.10.009 & 0.92 & Dynamical Systems \\
18892 & 10.1287/msom.1080.0228 10.1287/msom.1060.0190 & 0.72 & Operations Research\\ 
18918 & 10.1007/s11128-010-0177-y 10.1007/s11128-013-0567-z & 0.83 & Image Processing\\ 
18947 & 10.1109/TASE.2011.2160452 10.1109/TASE.2011.2178023 & 0.77 & Robotics\\ 
18956 & 10.1109/ICCV.2017.32 10.1109/ICCV.2017.31 & 0.84 & Computer Vision\\ 
18972 & 10.1109/ASE.2013.6693094 10.1145/2568225.2568254 & 0.69 & Software \\ 
18974 & 10.1016/j.amc.2009.03.023 10.1016/j.amc.2010.07.064 & 0.54 & Applied Mathematics\\ 
18983 & 10.1137/110848864 10.1137/110848876 & 0.66 & Optimization\\ 
18997 & 10.1504/IJMIC.2014.065339 10.1504/IJMIC.2015.068871 & 0.91 & Chaotic Systems \\ 
19008 & 10.1016/j.chb.2008.12.013 10.1016/j.chb.2008.12.012 & 0.63 & Hum Comp Interaction\\ 
19009 & 10.1109/AVSS.2017.8078491 10.1109/ICCV.2017.206 & 0.78 & Artificial Intelligence\\ 
19013 & 10.1145/2508859.2516668 10.1007/978-3-642-42045-0\_15 & 0.88 & Security\\ 
19026 & 10.1145/2541940.2541942 10.1109/HPCA.2014.6835965 & 0.81 & Architecture\\ 
19027 & 10.1016/S0305-0548(03)00250-8 10.1016/j.ejor.2006.03.013 & 0.82 & Graph algorithms\\ 
19030 & 10.1109/ICDE.2008.4497474 10.14778/1687627.1687666 & 0.69 & Databases \\ 
19031 & 10.1109/TMM.2005.843347 10.1109/TCE.2005.1405724 & 0.55 & Networks\\ 
19033 & 10.1109/TIFS.2014.2327757 10.1109/TCYB.2014.2376934 & 0.78 & Facial Recognition\\ 
19034 & 10.1016/j.neuroimage.2013.05.018 10.1016/j.neuroimage.2014.06.016 & 0.63 & Neurology\\ 
19035 & 10.1002/int.21933 10.1002/int.21927 & 1.09 & Intelligent Systems\\ 
19036 & 10.1006/jsco.2000.0402 10.1006/jsco.2000.0403 & 0.73 & Comp Geometry\\ 
\noalign{\smallskip}\hline
\end{tabular}
\end{table}
\newpage
